On the problem of repeated supervised learning

### 2023-03-28 paper/main.tex
article completion

    \begin{equation}
        \label{R}
        \mathbf{R} := \left\{f : \mathbb{R}^n \rightarrow \mathbb{R}_+ ~\text{and}~ \int\limits_{\mathbb{R}^n}f(x)dx = 1\right\},
    \end{equation}

@@

    A range of problems, that we are considering in this paper are conditions under which $\text{D}$ translates $\mathbf{R}$ into $\mathbf{R}$, conditions when $\text{D}$ is a contraction mapping and its fixed point existence conditions.

    Main contributions ...

@@

    An important contribution of this paper is that ...

@@

    In this paper we consider the following research questions:

    \begin{enumerate}
        \item[\textbf{1.}]  What are the conditions so that the mapping $\text{D}$ becomes a transformation $\mathbf{R}$ into $\mathbf{R}$?

        \item[\textbf{2.}] Under what conditions the mapping $\text{D}$ is a contraction mapping?

        \item[\textbf{3.}] Under what conditions the mapping $\text{D}$ has a fixed point?
    \end{enumerate}

    Let's analyze the importance of each question. We consider $\text{D}$ as operator of changing distribution of our dataset, so it has to transform $\mathbf{R}$ into $\mathbf{R}$. If $\text{D}$ is a contraction mapping, then for any starting function $f_0$ after a sufficiently large T, the result of our algorithm after $T$ steps would be $D^T(f_0) = f^*$, where $f^*$ is a fixed point, so after $T$ steps metrics of our algorithm cannot be improved. If $\text{D}$ has a fixed point $f$, when if $f_0 = f$ and we cannot improve our metrics as well.

@@

@@

    \subsection{Experiment Design} \label{design}
        We consider such a formulation of the problem of repeated supervised learning: there is an original data $\textbf{X}$ and a vector of target variables $\mathbf{y}$. Initially, at round $r = 0$, we select $30\%$ of the original dataset on which our model $f(\mathbf{x}, \theta^0)$ is trained with $80\%$ train size.

        Then at each step $t$ we randomly select an element $(\mathbf{x^i}, y_i)$ -- the features and target variable of the $i$-th object from the original dataset, and this element does not lie in the initial dataset. Next, we get the prediction $y'_i$ of our model on the element $(\mathbf{x^i}, y_i)$: $y_i'=f(\mathbf{x^i}, \mathbf{\theta}^r)$ and sample $z_i \sim \mathcal{N}(y_i', s \cdot \sigma^2)$, where $s$ is an experiment parameter that indicates adherence. Then we remove 1 element from the active dataset and, with the probability of $p$, we add $(\mathbf{x^i}, z_i)$ to it, and with the probability $(1-p)$ we add the element $(\mathbf{x^i}, y_i)$. We carry out this procedure until we run out of elements that were not initially included in our dataset, so, we make a total of $0.7 \cdot n$ steps, where $n$ is the number of objects in $\textbf{X}$.

        After each $T$ steps round $r$ is increasing: $r = r+1$ and the machine learning model $f(x, \theta)$ is retrained with $80\%$ train size on active dataset.

        The size of active dataset will always be $0.3 \cdot n$, because on each step we remove 1 element and add 1 element to active data. This experiment design is similar to \cite{khritankov2021hidden} were author was detecting hidden feedback loops in machine learning systems.

        In this paper we consider linear model, that is solved as Ridge with mean squared error loss function. In the formulation of our experiment, we consider $\mathbf{R}$ \eqref{R} as space of density functions of random vectors $(\mathbf{x^i}, y_i)$, where $\mathbf{x^i} \in \textbf{X}$ and $y_i \in \textbf{y}$ -- the features and target variable of the $i$-th object. The operator $\text{D}$ transforms $\mathbf{R}$ at each step $t$, and the distribution of features does not change, because at each step we take new $\mathbf{x^i}_t$ from the original set of features $\textbf{X}$, so only the distribution of the target variable changes. 
        If in some step the model starts to give good predictions on the training and test samples, then, in the probabilistic formulation, this means that the density function of the distribution of the object-sign vectors becomes similar to the delta function, as the components of the random vector $(\mathbf{x^i}, y_i)$ become linearly dependent.

@@

    \subsection{Experiment results} \label{res}
        First we consider $\mathbf{X}$ and $\mathbf{y}$ as a regression problem, so $\mathbf{y} = \mathbf{X}  \cdot \mathbf{\theta}$ for some vector $\mathbf{\theta}$. To complicate the model, a normally distributed noise was added to the data. This 3-dimensional plot will illustrate our results

>>>>

    \begin{equation}
        \label{R}
        \mathbf{R} := \left\{f : \mathbb{R}^n \rightarrow \mathbb{R}_+ ~\text{and}~ \int\limits_{\mathbb{R}^n}f(x)dx = 1\right\}
    \end{equation}

@@

    In this paper we propose a mathematical model of the process of repeated learning that is new to the literature. We find sufficient conditions for the operator $\text{D}$ to translate $\mathbf{R}$ into $\mathbf{R}$. We find conditions for our data density functions to tend in a weak sense to a delta function under the operator $\text{D}$. We also find sufficient conditions for the operator $\text{D}$ to be non-contraction in the norm $\|\cdot \|_q$. The importance of these properties is explained in detail in Section \ref{Main_results}.

@@

    An important contribution of this paper is that  we built a mathematical model for continuous machine learning problem, which is novel to the literature.

@@

    In the next section, we provide several conditions on the operator $\text{D}$, in order for it to translate $\mathbf{R}$ into $\mathbf{R}$. It is important, because we consider $\text{D}$ as operator of changing distribution of our dataset, so it has to transform $\mathbf{R}$ into $\mathbf{R}$.

    We also look at the conditions under which the data density functions will tend to a delta function in the weak sense, i.e. $\text{D}^t(f_0)(x) \underset{t \to +\infty}{\longrightarrow} \delta(x)$. 
    This is an important property, because we it can be used to understand when repeated machine learning improves our metrics. Why this is so will be discussed in detail in Section \ref{Main_results}.

    Another contribution of our paper is a condition, under that operator $\text{D}$ wouldn't be a contraction mapping in any norm $\|\cdot\|_q$, i.e. there always would be function $f \in \mathbf{R}$ such that $\|\text{D}(f)\|_q \geq \|f\|_q$. This is also important property, because if $\text{D}$ is a contraction mapping, then it converge any start density distribution $f_0$ to function that is equal to zero in almost every point $x \in \mathbb{R}^n$, i.e. continuous algorithm transform our data to uniform noise.

@@

    \textbf{Notation:} In this paper we will use the common notations: the $L_1(\mathbb{R}^n)$-norm of function $f$:

    \begin{equation*}
        \|f\|_1 := \int\limits_{\mathbb{R}^n} |f(x)| dx \quad \text{ and } \quad L_1(\mathbb{R}^n) := \left\{f : \mathbb{R}^n \to \mathbb{R} ~|~ \|f\|_1 < + \infty\right\}
    \end{equation*}

    The $L_{\infty}(\mathbb{R}^n)$-norm of function $f$:

    \begin{equation*}
        \|f\|_{\infty} := \underset{x \in \mathbb{R}^n}{\text{esssup}}\{f(x)\} := \inf\{C \geq 0 ~|~ |f(x)| \leq C \text{ for almost every } x \in \mathbb{R}^n\} \quad \text{ and } \quad L_{\infty}(\mathbb{R}^n) := \left\{f : \mathbb{R}^n \to \mathbb{R} ~|~ \|f\|_{\infty} < + \infty\right\}
    \end{equation*}

    \begin{theorem}[Fact] \label{based}
        If $f: \mathbb{R}^n \to \mathbb{R}$ such that $f(x) \geq 0$ for almost every $x \in \mathbb{R}^n$ and $\|f\|_1 = \int\limits_{\mathbb{R}^n} f(x) dx = 1$, then there exists a random vector $\mathbf{\xi}$, for which $f$ will be a density distribution function.
    \end{theorem}

    Exactly on the basis of Theorem \ref{based} we define $\mathbf{R}$ \eqref{R} in this way.

    \begin{theorem}[Assumptions for $\text{D}: \mathbf{R} \to \mathbf{R}$] \label{R_to_R}
        If $\|\text{D}\|_1 = 1, \forall f \in \mathbf{R} \hookrightarrow \text{D}(f)(x) \geq 0$ for almost every $x \in \mathbb{R}^n$, and exists $\text{D}^{-1}$ such that $\|\text{D}^{-1}\|_1 \leq 1$, then $\text{D} : \mathbf{R} \to \mathbf{R}$.
    \end{theorem}
    \begin{proof}
        To begin with, let us note that if $\text{D}: \mathbf{R} \to \mathbf{R}$, then $\|\text{D}\|_1 = 1$, because by definition of operator norm:

        \begin{equation*}
            \|\text{D}\|_1 \overset{def}{=} \underset{\|f\|_1 = 1}{\sup}\left\{\|\text{D}(f)\|_1\right\}
        \end{equation*}

        And if $f$ such that $\|f\|_1 = 1$ then $|f| \in \mathbf{R}$ and, because $\text{D}: \mathbf{R} \to \mathbf{R}$, $\|\text{D}(|f|)\|_1 = 1$. But $\|\text{D}\|_1 = 1$ is only a necessary but not a sufficient condition.

        If $\|\text{D}\|_1 = 1$, then $\forall f \in \mathbf{R} \hookrightarrow \text{D}(f) \leq 1$. If $\exists f_0 \in \mathbf{R}$ such that $\|\text{D}(f_0)\|_1 < 1$, then we get a contradiction because

        \begin{equation*}
            \|D^{-1}\|_1 \overset{def}{=} \underset{\|f\|_1 \neq 0}{\sup}\left\{\dfrac{\|\text{D}^{-1}(f)\|_1}{\|f\|_1}\right\} 
            \geq \left[f_1 = \text{D}(f_0)\right] \geq
            \dfrac{\|\text{D}^{-1}(f_1)\|_1}{\|f_1\|_1} = 
            \dfrac{\|\text{D}^{-1}(\text{D}(f_0))\|_1}{\|\text{D}(f_0)\|_1} = 
            \dfrac{\|f_0\|_1}{\|\text{D}(f_0)\|_1} = \dfrac{1}{\|\text{D}(f_0)\|_1} > 1
        \end{equation*}

        But we assume that $\|\text{D}^{-1}\|_1 \leq 1$. So $\forall f \in \mathbf{R} \hookrightarrow \|\text{D}(f)\|_1 = 1$.

        But according to Theorem \ref{R_to_R} to $\text{D} : \mathbf{R} \to \mathbf{R}$ we also need second assumption: $\forall f \in \mathbf{R} \hookrightarrow \text{D}(f)(x) \geq 0$ for almost every $x \in \mathbb{R}^n$.

    \end{proof}

    \textbf{Discussion of Theorem \ref{R_to_R}}

    In experiments it often difficult to calculate $\text{D}^{-1}$ and especially it's norm, so we make a different assumptions. We consider $\text{D}$ as algorithm for training a model, forming predictions on a test sample and mixing predictions into a training sample. Distribution of our data is approximating by empirical distribution function \cite{dvoretzky1956asymptotic} as follows (for $n=1$):

    \begin{equation}\label{F_approx}
        \hat{F}_N(x) := \dfrac{\text{number of elements in sample} \leq x}{N} = \dfrac{1}{N}\sum\limits_{i=1}^N \textbf{1}_{X_i \leq x},
    \end{equation}

    where $X_i$ are elements of sample. We assume that $(X_1, X_2, X_3, ... , X_N)$ are independent, identically distributed real random variables with the common cumulative distribution function $F(x)$. If this assumption is fulfilled, then the DKW inequality is satisfied:

    \begin{equation}\label{DKW}
        \mathbb{P}\left\{\underset{x \in \mathbb{R}}{\sup}\left|\hat{F}_N(x) - F(x)\right| > \varepsilon \right\} \leq C e^{-2N\varepsilon^2} \quad 
        \forall \varepsilon > 0
    \end{equation}

    And we can build interval that contains the true CDF of our data $F(x)$, with probability $1 - \alpha$ as

    \begin{equation}\label{inter}
        \hat{F}_N(x) - \varepsilon \leq F(x) \leq \hat{F}_N(x) + \varepsilon, ~~ \text{ where } ~~ \varepsilon = \sqrt{\dfrac{\ln(2/\alpha)}{2N}}
    \end{equation}

    In this case operator $\text{D}$ transoms our data, i.e. translates one empirical distribution function to another. So $\text{D} : \mathbf{R} \to \mathbf{R}$ by constructing our experiment.

    \begin{theorem}[Limit in a weak sense to $\delta$ function] \label{delta}
        If $f_t : \mathbb{R} \to \mathbb{R}$ such that $\forall t \in \mathbb{N} \hookrightarrow  \|f_t\|_1 = 1$, $f_t(x) \geq 0$ in almost every point $x \in \mathbb{R}$ and
        \begin{equation} \label{psi_and_g}
            \exists \psi : \mathbb{N} \to \mathbb{R} : \psi(t) \underset{t \to +\infty}{\longrightarrow} +\infty ~\text{ and }~
            \exists g \in L_1(\mathbb{R}) ~\text{ such that }~ \forall t \in \mathbb{N} ~\forall y \in \mathbb{R} \hookrightarrow f_t\left(\dfrac{y}{\psi(t)}\right) \leq \psi(t) \cdot |g(y)|
        \end{equation}

        Then $f_t(x) \underset{t \to \infty}{\longrightarrow} \delta(x)$ in a weak sense, i.e.

        \begin{equation}
            \underset{t \to +\infty}{\lim}\left(\int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx\right) = \phi(0),
        \end{equation}

        where $\phi$ is continuous function with compact support
    \end{theorem}

    \begin{proof}
        Assume a notation $I_t = \int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx$. Then it's fulfilled that

        \begin{equation*}
            I_t - \phi(0) = \int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx - \phi(0) \cdot \int\limits_{-\infty}^{+\infty} f_t(x) dx = \int\limits_{-\infty}^{+\infty} f_t(x) \cdot [\phi(x) - \phi(0)] dx
        \end{equation*}

        The first equation is fulfilled because $\|f_t\| = 1$.

        Replacing the variable $y = \psi(t) \cdot x, dy = \psi(t) \cdot dx$ we get

        \begin{equation} \label{tmp2}
            I_t - \phi(0) = \dfrac{1}{\psi(t)} \cdot \int\limits_{-\infty}^{+\infty} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation}

        Split the integral \eqref{tmp2} into 3 parts:

        \begin{equation*}
            I_1 := \dfrac{1}{\psi(t)} \cdot \int\limits_{-\infty}^{-A} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}
        \begin{equation*}
            I_2 := \dfrac{1}{\psi(t)} \cdot \int\limits_{-A}^{A} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}
        \begin{equation*}
            I_3 := \dfrac{1}{\psi(t)} \cdot \int\limits_{A}^{+\infty} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}

        Consider the integrals $I_1$. $\phi$ is continuous function with compact support, so $\phi$ is bounded by some constant $M$, i.e. $\forall x \in \mathbb{R} \hookrightarrow |\phi(x)| \leq M$, so

        \begin{equation*}
            \left|I_1\right| \leq \int\limits_{-\infty}^{-A} 2 M \cdot \dfrac{1}{\psi(t)} f_t\left(\frac{y}{\psi(t)}\right) dy \leq \left[\eqref{psi_and_g}\right]
            \leq \int\limits_{-\infty}^{-A} 2 M \cdot |g(y)| dy 
        \end{equation*}

        Since $g \in L_1(\mathbb{R})$, there exists some constant $A > 0$ such that $\left|I_1\right| \leq \varepsilon$. Similarly, it can be shown that $\left|I_3\right| \leq \varepsilon$.

        Consider the integral $I_2$. $\phi$ is continuous and $\psi \to +\infty$ \eqref{psi_and_g}, so $\exists~ T \in \mathbb{N}$ such that $\forall y \in [-A; A]$ $\forall t \geq T$ it is fulfilled that

        \begin{equation*}
            \left|\frac{y}{\psi(t)} - 0\right| \leq \delta ~\text{ and so }~ \left|\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right| \leq \varepsilon
        \end{equation*}

        So

        \begin{equation*}
            \left|I_2\right| \leq \varepsilon \cdot \int\limits_{-A}^{A} \dfrac{1}{\psi(t)} f_t\left(\frac{y}{\psi(t)}\right) dy = 
            \varepsilon \cdot \int\limits_{-A/\psi(t)}^{A/\psi(t)} f_t\left(x\right) dx \leq \varepsilon \cdot \int\limits_{-\infty}^{+\infty} f_t\left(x\right) dx = \varepsilon
        \end{equation*}

        Finally we get

        \begin{equation*}
            \forall \varepsilon > 0 ~~ \exists~ T \in \mathbb{N} : \forall t \geq T \hookrightarrow \left|I_t - \phi(0)\right| \leq 
            \left|I_1\right| + \left|I_2\right| + \left|I_3\right| \leq 3\varepsilon
        \end{equation*}

        So $f_t(x) \underset{t \to \infty}{\longrightarrow} \delta(x)$ in a weak sense.

    \end{proof}

    \textbf{Discussion of Theorem \ref{delta}}

    If in some step $t$ the model $H(x, \theta, t)$ starts to give good predictions on the training and test samples, then, in the probabilistic formulation, this means that the density function of the distribution of the object-sign vectors becomes similar to the delta function, as the components of the random vector $(\mathbf{x^i}, y_i) \subset (\mathbf{X}, \mathbf{y})$ become linearly dependent. Based on these considerations, we can compare each operator $\text{D}$ to an operator $\widetilde{\text{D}}$, where $\widetilde{\text{D}}$ transforms density distribution functions of random variables $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$, where $\mathbf{y_{\text{pred}}} = H(\mathbf{X}, \theta, t)$. Then we can apply Theorem \ref{delta} to understand, when $\widetilde{\text{D}}(f_0)(x) \underset{t \to \infty}{\longrightarrow} \delta(x)$.

    If $\widetilde{\text{D}}$ has a fixed point, then the distribution of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ does not change, so algorithm of repeated learning does not improves quality metrics, but in this case operator $\text{D}$ can still have no fixed point. That's why we consider formulas \eqref{DKW} and \eqref{inter} only for $n=1$, because in our experiments we will detect a moment, when MSE error of our algorithm will stop decreasing, so the distribution of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ will be a fixed point of $\widetilde{\text{D}}$.

    Let's analyze formula \eqref{psi_and_g}.

    If we take $x = \psi(t) \cdot y$ then \eqref{psi_and_g} takes form

    \begin{equation} \label{g_1}
            \exists g \in L_1(\mathbb{R}) ~\text{ such that }~ \forall t \in \mathbb{N} ~\forall x \in \mathbb{R} \hookrightarrow f_t\left(x\right) \leq \psi(t) \cdot |g(x \cdot \psi(t))|
    \end{equation}

    If $x \neq 0$ then $f_t(x) \underset{t \to \infty}{\longrightarrow} 0$, because $g_1(x) := \psi(t) \cdot |g(x \cdot \psi(t))| \in L_1(\mathbb{R})$ since

    \begin{equation*}
        \int\limits_{-\infty}^{+\infty} g_1(x) dx = \int\limits_{-\infty}^{+\infty} \psi(t) \cdot |g(x \cdot \psi(t))| dx = \int\limits_{-\infty}^{+\infty} g_1(z) dz < +\infty
    \end{equation*}

    And so, if $t \to \infty$, then $z := x \cdot \psi(t) \to +\infty$ and $g_1(z) \to 0$, because $g_1 \in L_1(\mathbb{R})$. So, if $x \neq 0$ then $f_t(x) \underset{t \to \infty}{\longrightarrow} 0$.

    Since $\forall t \in \mathbb{R} \hookrightarrow \|f_t\|_1 = 1$, then $f_t(0) \to +\infty$.

    If we substitute $x = 0$ in the \eqref{g_1} then we get $f_t(0) \leq \psi(t) \cdot |g(0)|$, so we can take

    \begin{equation}
        \psi(t) = \dfrac{f_t(0)}{|g(0)|}
    \end{equation}

    In our experiments we will measure $f_t(0)$ and $\int\limits_{-\kappa}^{\kappa}f_t(x)dx = \hat{F}_t(\kappa) - \hat{F}_t(-\kappa)$, where $\kappa$ is sufficiently small. So if $f_t(0) \to +\infty$ and $1 \in I_{\varepsilon}(\hat{F}_t(\kappa) - \hat{F}_t(-\kappa))$, where $I_{\varepsilon}$ is confidence interval from \eqref{inter}, then we can say that operator $\widetilde{\text{D}}$ converge empirical distribution functions $f_t(x)$ of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ to delta-function.

    Now we can see importance of equations \eqref{DKW} and \eqref{inter}. If empirical density distributions functions $\widetilde{\text{D}}$ converge to delta functions then according to \eqref{DKW} and \eqref{inter} we can consider, that true density distribution function $F(x)$ converge to $\delta(x)$, because from \eqref{inter} we have

    \begin{equation*}
        \hat{F}_t(x) - \varepsilon \leq F(x) \leq \hat{F}_t(x) + \varepsilon, ~~ \text{ where } ~~ \varepsilon = \sqrt{\dfrac{\ln(2/\alpha)}{2N}} ~\text{ and }~ \hat{F}_t(x) \text{ is empirical CFD on step } t
    \end{equation*}

    So if $\hat{F}_t(x) \to F_{\delta}(x) = \text{sign}(x)$, then $F(x) \to F_{\delta}(x)$

    Important example of operator $\text{D}$ that translates any function from $\mathbf{R}$ into a $\delta$ function is as follows

    \begin{equation} \label{example1}
        \text{D}^t(f_0)(x) = t \cdot f_0(t \cdot x)
    \end{equation}

    Here we take $g(x) = f_0(x)$ and $\psi(t) = t$.

    In Figure \ref{example1_fig} shown how this operator translates density functions of normal distribution $\mathcal{N}(0, 5)$ and continuous uniform distribution $\mathcal{U}[-2.5, 2.5]$ with $t \to +\infty$:

    \begin{figure}[h!]
        \centering
        \includegraphics[width = 0.99\linewidth]{pictures/example1.png}
        \caption{Illustration of weak limit to $\delta$ function. $\mathcal{N}(0, 5)$ left, $\mathcal{U}[-2.5, 2.5]$ right.}
        \label{example1_fig}
    \end{figure}

    \begin{theorem}[Inequality on $\|\text{D}\|_q$] \label{ineq_q}
        Consider 
        \begin{equation}\label{f_R}
            f_R(x) = \dfrac{1}{\text{vol}(\mathbb{B}(R))} \cdot \textbf{1}_{\mathbb{B}(R)}(x),
        \end{equation}

        where $\mathbb{B}(R) := \{x \in \mathbb{R}^n ~|~ \|x\|_2 \leq 1\}$ and $\text{vol}(\mathbb{B}(R))$ - the volume of this Euclidean ball.

        Then for all $R > 0$ and for all $1 \leq q \leq +\infty$ such that $\text{D}(f_R) \in L_q(\mathbb{R}^n)$ is fulfilled that

        \begin{equation}\label{int_f_R}
            \|\text{D}\|_q \geq \int\limits_{\mathbb{B}(R)} \text{D}(f_R)(x)dx
        \end{equation}
    \end{theorem}

    \begin{proof}
        First of all let's calculate $\|f_R\|_p$:

        \begin{equation*}
            \|f_R\|_p = \left(\int\limits_{-\infty}^{+\infty}\left(\dfrac{1}{\text{vol}(\mathbb{B}(R))}\right)^p \cdot \textbf{1}_{\mathbb{B}(R)}(x) dx\right)^{1/p} = \dfrac{1}{\text{vol}(\mathbb{B}(R))} \cdot \text{vol}(\mathbb{B}(R))^{1/p} = \text{vol}(\mathbb{B}(R))^{-1 + 1/p}
        \end{equation*}

        So, $f_R \in \in L_p(\mathbb{R}^n)$ for all $R > 0$ and $1 \leq p \leq +\infty$.

        Now write out a Herder's inequality. For $q$ such that $\frac{1}{p} + \frac{1}{q} = 1$ is fulfilled that

        \begin{equation*}
            \|f_R\|_p \cdot \|\text{D}(f_R)\|_q \geq \|f_R \cdot \text{D}(f_R)\|_1
        \end{equation*}

        Using common inequality on operators norm $\|\text{D}(f)\|_q \leq \|\text{D}\|_q \cdot \|f\|_q ~\forall f \in L_q(\mathbb{R}^n)$ we get

        \begin{equation*}
            \|f_R\|_p \|f_R\|_q \cdot \|\text{D}\|_q \geq \|f_R \cdot \text{D}(f_R)\|_1
        \end{equation*}

        Since $\|f_R\|_p \|f_R\|_q = \text{vol}(\mathbb{B}(R))^{-1 + 1/p} \cdot \text{vol}(\mathbb{B}(R))^{-1 + 1/q} = \text{vol}(\mathbb{B}(R))^{-2 + 1/p + 1/q} = \text{vol}(\mathbb{B}(R))^{-1}$ we get:

        \begin{equation*}
            \|\text{D}\|_q \geq \text{vol}(\mathbb{B}(R)) \cdot \|f_R \cdot \text{D}(f_R)\|_1
        \end{equation*}

        Let's look at $\|f_R \cdot \text{D}(f_R)\|_1$ in more detail:

        \begin{equation*}
            \|f_R \cdot \text{D}(f_R)\|_1 = \int\limits_{-\infty}^{+\infty} \text{D}(f_R)(x) \cdot \dfrac{1}{\text{vol}(\mathbb{B}(R))} \textbf{1}_{\mathbb{B}(R)}(x) dx = \dfrac{1}{\text{vol}(\mathbb{B}(R))} \int\limits_{\mathbb{B}(R)} \text{D}(f_R)(x)dx
        \end{equation*}

        Finally, we get the desired inequality

        \begin{equation*}
            \|\text{D}\|_q \geq \int\limits_{\mathbb{B}(R)} \text{D}(f_R)(x)dx
        \end{equation*}
    \end{proof}

    \textbf{Discussion of Theorem \ref{ineq_q}}

    To begin with, note that the result of Theorem \ref{ineq_q} does not depend in any way on whether $\text{D}$ converts $\mathbf{R}$ to $\mathbf{R}$, it is a consequence of Herder's inequality.

    If you look at it from a probabilistic point of view (so now $\text{D} : \mathbf{R} \to \mathbf{R}$), then $f_R$ is a distribution density function of vectors uniformly distributed on a Euclidean sphere $\mathbb{B}(R)$. So, $\int\limits_{\mathbb{B}(R)} \text{D}(f_R)(x)dx \leq 1$, i.e. from Theorem \ref{ineq_q} we can only get that $\|\text{D}\|_q \geq 1$. But if $\|\text{D}\|_q \geq 1$, so $\text{D}$ wouldn't be a contraction mapping in $\|\cdot\|_q$, because there always would be function $f \in \mathbf{R}$ such that $\|\text{D}(f)\|_q \geq \|f\|_q$.

    In our experiments we will measure $\hat{F}_t(R)$ and $\hat{F}_t(-R)$. If $1 \in I_{\varepsilon}(\hat{F}_t(R))$ and $0 \in I_{\varepsilon}(\hat{F}_t(-R))$, where $I_{\varepsilon}$ is confidence interval from \eqref{inter}, then we can say that operator $\text{D}$ can't be a contraction mapping.

@@

    \subsection{Experiment Design} \label{design}
        We consider such a formulation of the problem of repeated supervised learning: there is an original data $\textbf{X}$ and a vector of target variables $\mathbf{y}$. Initially, at round $r = 0$, we select $30\%$ of the original dataset on which our model $H(\mathbf{x}, \theta^0, 0)$ is trained with $80\%$ train size.

        Then at each step $t$ we randomly select an element $(\mathbf{x^i}, y_i)$ -- the features and target variable of the $i$-th object from the original dataset, and this element does not lie in the initial dataset. Next, we get the prediction $y'_i$ of our model on the element $(\mathbf{x^i}, y_i)$: $y_i'=H(\mathbf{x^i}, \mathbf{\theta}^r, t)$ and sample $z_i \sim \mathcal{N}(y_i', s \cdot \sigma^2)$, where $s$ is an experiment parameter that indicates adherence and $\sigma$ is the model's mean squared error on held-out data. Then we remove 1 element from the active dataset and, with the probability of $p$, we add $(\mathbf{x^i}, z_i)$ to it, and with the probability $(1-p)$ we add the element $(\mathbf{x^i}, y_i)$. We carry out this procedure until we run out of elements that were not initially included in our dataset, so, we make a total of $0.7 \cdot n$ steps, where $n$ is the number of objects in $\textbf{X}$.

        After each $T$ steps round $r$ is increasing: $r = r+1$ and the machine learning model $H(x, \theta^r, t)$ is retrained with $80\%$ train size on active dataset.

        The size of active dataset will always be $0.3 \cdot n$, because on each step we remove 1 element and add 1 element to active data. This experiment design is similar to \cite{khritankov2021hidden} were author was detecting hidden feedback loops in machine learning systems. The scheme of the experiment is shown in Figure \ref{ex_set}.

        \begin{figure}[h!]
            \centering
            \includegraphics[scale = 0.5]{pictures/Experiment_setup.png}
            \caption{Experiment setup.}
            \label{ex_set}
        \end{figure}

        In this paper we consider linear model, that is solved as Ridge with mean squared error loss function. In the formulation of our experiment, we consider $\mathbf{R}$ \eqref{R} as space of density functions of random vectors $(\mathbf{x^i}, y_i)$, where $\mathbf{x^i} \in \textbf{X}$ and $y_i \in \textbf{y}$ -- the features and target variable of the $i$-th object. The operator $\text{D}$ transforms $\mathbf{R}$ at each step $t$, and the distribution of features does not change, because at each step we take new $\mathbf{x^i}_t$ from the original set of features $\textbf{X}$, so only the distribution of the target variable changes. So we can use results from Theorem \ref{delta}.

\newpage

    \subsection{Experiment results} \label{res}

@@

        First we consider $\mathbf{X}$ and $\mathbf{y}$ as a regression problem, so $\mathbf{y} = \mathbf{X}  \cdot \mathbf{\theta}$ for some vector $\mathbf{\theta}$. To complicate the model, a normally distributed noise was added to the data. In Figure \ref{3D} there are five 3-dimensional plots for different noise in original data, where on $X$-axis is plotted \textit{usage}, on $Y$-axis is \textit{adherence} and on $Z$ is deviation of $\mathbf{y} - \mathbf{y_{\text{pred}}}$ for test data samples.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.33\linewidth]{pictures/3D_plot_0.1.png}
            \includegraphics[width=0.33\linewidth]{pictures/3D_plot_0.3.png}
            \includegraphics[width=0.33\linewidth]{pictures/3D_plot_1.png}

            \includegraphics[width=0.33\linewidth]{pictures/3D_plot_3.png}
            \includegraphics[width=0.33\linewidth]{pictures/3D_plot_10.png}

            \caption{3D Graphics for different  noise in original data}
            \label{3D}
        \end{figure}

        As we can see, the picture for the graphs at noise= $0.3, 1, 3$ and $10$ are very similar to each other, they differ only in the deviation values. This is due to the fact that for higher values of noise it becomes more profitable to add new data than to take the original data, since the noises in them are smaller, that is, with increasing \textit{usage} the deviation will fall.

\newpage
